---
title: "STAT/MATH 495: Problem Set 10"
author: "Tasheena Narraidoo"
date: "2017-11-28"
output:
  html_document:
    toc: true
    toc_float: true
    toc_depth: 2
    collapsed: false
    smooth_scroll: false
    df_print: kable
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.width=8, fig.height=4.5, message=FALSE, warning = FALSE)
```



# Collaboration

Please indicate who you collaborated with on this assignment: 



# Setup

```{r, echo=FALSE}
#load libraries
library(tidyverse)
library(broom)
library(glmnet)
library(Metrics)
```


```{r}
train <- read_csv("data/train.csv")
test <- read_csv("data/test.csv")
sample_submission <- read_csv("data/sample_submission.csv")

# Only use 150 observations to train model!
set.seed(76)
train <- train %>% 
  mutate(log_price_doc = log(price_doc)) %>% 
  sample_n(150)

# Need "dummy" outcome variable to make model.matrix() code below work
test <- test %>% 
  mutate(log_price_doc=1) 

# Model formula
model_formula <- as.formula("log_price_doc ~ full_sq + area_m + raion_popul + green_zone_part + indust_part + children_preschool + preschool_education_centers_raion + children_school + school_education_centers_raion + school_education_centers_top_20_raion + healthcare_centers_raion + university_top_20_raion + sport_objects_raion + additional_education_raion + culture_objects_top_25 + culture_objects_top_25_raion + shopping_centers_raion + office_raion + thermal_power_plant_raion + incineration_raion + oil_chemistry_raion + radiation_raion + railroad_terminal_raion + big_market_raion + nuclear_reactor_raion + detention_facility_raion + full_all + male_f + female_f + young_all + young_male + young_female + work_all + work_male + work_female + ekder_all + ekder_male + ekder_female + ID_metro + metro_min_avto + metro_km_avto + kindergarten_km + school_km + park_km + green_zone_km + industrial_km + water_treatment_km + cemetery_km + incineration_km + railroad_station_avto_km + railroad_station_avto_min + ID_railroad_station_avto + public_transport_station_km + public_transport_station_min_walk + water_km + water_1line + mkad_km + ttk_km + sadovoe_km + bulvar_ring_km + kremlin_km + big_road1_km + ID_big_road1 + big_road1_1line + big_road2_km + ID_big_road2 + railroad_km + railroad_1line + zd_vokzaly_avto_km + ID_railroad_terminal + bus_terminal_avto_km + ID_bus_terminal + oil_chemistry_km + nuclear_reactor_km + radiation_km + power_transmission_line_km + thermal_power_plant_km + ts_km + big_market_km + market_shop_km + fitness_km + swim_pool_km + ice_rink_km + stadium_km + basketball_km + hospice_morgue_km + detention_facility_km + public_healthcare_km + university_km + workplaces_km + shopping_centers_km + office_km + additional_education_km + preschool_km + big_church_km + church_synagogue_km + mosque_km + theater_km + museum_km + exhibition_km + catering_km + green_part_500 + prom_part_500 + office_count_500 + office_sqm_500 + trc_count_500 + trc_sqm_500") 

# Define predictor matrices
predictor_matrix_train <- model.matrix(model_formula, data = train)[, -1]
predictor_matrix_test <- model.matrix(model_formula, data = test)[, -1]
```


# Do work and create submission files:

```{r, message=FALSE}
# 1. linear model

# Fit linear model
lin_mod <- lm(model_formula, data=train)
# Make prediction with test set
lin_pred <- exp(predict(lin_mod, test))
# prepare submission file
lin_submission <- sample_submission %>% 
  mutate(price_doc = lin_pred) %>% 
  select(id, price_doc)
write.csv(lin_submission, "linear_submission.csv")  

# compute rmsle for linear model
# predict on train data
lin_mod_train <- predict(lin_mod, train)
# calc rmsle using Metrics library
rmsle(train$log_price_doc, lin_mod_train) # 0.01577062

```

```{r, message=FALSE}
# 2. LASSO model

# perform lasso cv to find optimal lambda
lasso_cv <- cv.glmnet(x=predictor_matrix_train, y=train$log_price_doc, alpha=1)
lambda_opti <- lasso_cv$lambda.min
# fit lasso with optimal lambda
lasso_mod <- glmnet(x=predictor_matrix_train, y=train$log_price_doc, alpha=1, lambda = lambda_opti)
# make prediction using test set
lasso_pred <- exp(predict(lasso_mod, predictor_matrix_test))
# prepare submission file
lasso_submission <- sample_submission %>% 
  mutate(price_doc = lasso_pred) %>% 
  select(id, price_doc)
write.csv(lasso_submission, "lasso_submission.csv")  

# compute rmsle for lasso model
# predict on train data
lasso_mod_train <- predict(lasso_mod, predictor_matrix_train)
# calc rmsle using Metrics library
rmsle(train$log_price_doc, lasso_mod_train) # 0.0284825

```

```{r}
# plot lasso
plot(lasso_cv)
abline(v=log(lasso_cv$lambda.min), col="blue")
```

From the above plot we can see the cross-validated curve (red-dotted line) and the blue vertical line corresponds to the minimum lambda value. The other dotted vertical line represents `lambda.1se` lambda value which gives the most regularized model which is within 1 standard error of the minimum lambda value.


# Scoreboard

Using the "scoring mechanism" for the Russian Housing Kaggle competition, fill
in these cells:


Method                | Training Score  | Kaggle Score
--------------------- | -------------   | -------------
lm                    | 0.0158          |  2.63
LASSO crossvalidated  | 0.0285          |  0.419
